Popularizing linear and nonlinear global approaches to hydrodynamic instabilities : A review and a simple implementation for the wake of a cylinder

### 2017-10-09 ARTICLE_STABFEM/ARTICLE_ASME.tex
minor fix

To build an optimal mesh for the base-flow calculation, the idea is to use as the objective function $u_h$ the solution ${\bf u}_b$ itself, as computed on a previous mesh.%The procedure thus ensures that the mesh will be locally adapted to the gradients of the base flow. 
The base flow is then recomputed on the adapted mesh, providing a better approximation of the solution. The procedure can be repeated a few steps to ensure a right convergence.
In our implementation, the whole process is performed using the Matlab driver
 {\em Freefem\_AdaptMesh.m}.

@@

@@

@@

We will now illustrate the applicati

@@

@@

%popular class of method are {\em power methods} which allow to compute a limited set

%based on the idea that the repeated application of an operator mostly amplifies the perturbations along the eigenspaces with largest eigenvalues.

%These de Andrea Fani

%\paragraph{Implementation and mesh adaptation for stability calculations}

Matlab driver...

%\paragraph{Sample results}

%\subsection{Illustration of the mesh adaptation for stability calculations}

%See listing....

%\subsection{Sample results for a cylinder}

@@

@@

\section{Detail of the weak formulation}
%The problem can be set into weak form by multiplying by multiplying Eq. \ref{Newton2} by a test function ${\bf v}$ and the associated divergence constraint by a test function $q$. After a few integration by parts we are led to :
\begin{eqnarray}
\label{NewtonWeak}
&\forall ({\bf v};q), \\
\displaystyle &\int \left[ {\bf v} \cdot {\cal C}( {\bf u}_b^g , \delta {\bf u}_b) +  \nabla  \cdot \delta {\bf u}_b q -\nabla  \cdot {\bf v} \delta p_b
+ \frac{2}{Re} {\bf D}(\delta {\bf u}_b) : {\bf D}({\bf v}) \right]
\nonumber
\\
\nonumber
\displaystyle & + \int \left[ {\bf v} \cdot ( {\bf u}_b^g \cdot \nabla {\bf u}_b^g) 
+ \nabla \cdot {\bf u}_b^g  q 
- \nabla \cdot {\bf v} p_b^g
+ \frac{2}{Re} {\bf D}({\bf u}_b^g) : {\bf D}({\bf v}) \right] = 0 
\end{eqnarray}

>>>>

@@

To build an optimal mesh for the base-flow calculation, the idea is to use as the objective function $u_h$ the solution ${\bf u}_b$ itself, as computed on a previous mesh.%The procedure thus ensures that the mesh will be locally adapted to the gradients of the base flow. 
The base flow is then recomputed on the adapted mesh, providing a better approximation of the solution. The procedure can be repeated a few steps to ensure a right convergence.

The mesh generated in the previous way may not be optimal for the stability calculations as the structure of the eigenmode may be more complex than that of the base flow. To remedy with this, the idea is to subsequently adapt the mesh to both the mesh flow and the results of the stability calculation. This is easily done with FreeFem, as the  AdaptMesh procedure can be used with several objective functions. We have experimented two different ideas. The first is to adapt the mesh to the base flow and the structure of the leading eigenmode. A second and smarter idea is to adapt the mesh to the base flow and the structural sensitivity. Since this quantity is directly linked to the sensitivity of the eigenvalues to perturbations of any kind (including spatial discretization errors), this procedure yields a lighter mesh while ensuring that the eigenvalue is effectively converged with respect to discretization errors. 
On the other hand, if one is interested to know the structure of the eigenmode (which may extend far away from the wavemaker region), it is more recommended to use mesh adaptation on the structure of the eigenmode.

In our implementation, the whole process is performed using the Matlab driver {\em Freefem\_AdaptMesh.m} ; the kind of adaptation (to base flow only, to eigenmode, or to sensitivity) is decided by the choice of parameters transmitted to this function.

@@

SECTION TO BE COMPLETED.

@@

Cylinder of unit diameter (radius $r=1/2$), velocity $U = 1$ at inlet.

Domain dimension $[-40,80]x[0,40]$. No-stress conditions at outlet AND LATERAL BOUNDARY.

@@

Lines 1-7 of the script.

@@

Figure \ref{fig:Eigenmode} : description of eigenmode.

Figure \ref{fig:SigmaOmega} : growth rates/Strouhal.

@@

SECTION TO BE COMPLETED

@@

\section{Detail of the weak formulation}
%The problem can be set into weak form by multiplying by multiplying Eq. \ref{Newton2} by a test function ${\bf v}$ and the associated divergence constraint by a test function $q$. After a few integration by parts we are led to :

SECTION TO BE COMPLETED.

We have to explain the integration by parts of the viscous term.

\begin{eqnarray}
\label{NewtonWeak}
&\forall ({\bf v};q), \\
\displaystyle &\int \left[ {\bf v} \cdot {\cal C}( {\bf u}_b^g , \delta {\bf u}_b) +  \nabla  \cdot \delta {\bf u}_b q -\nabla  \cdot {\bf v} \delta p_b
+ \frac{2}{Re} {\bf D}(\delta {\bf u}_b) : {\bf D}({\bf v}) \right]
\nonumber
\\
\nonumber
\displaystyle & + \int \left[ {\bf v} \cdot ( {\bf u}_b^g \cdot \nabla {\bf u}_b^g) 
+ \nabla \cdot {\bf u}_b^g  q 
- \nabla \cdot {\bf v} p_b^g
+ \frac{2}{Re} {\bf D}({\bf u}_b^g) : {\bf D}({\bf v}) \right] = 0 
\end{eqnarray}
